---
title: "Lab2"
author: "Joshua Wiseman"
date: "2023-01-27"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

 Working directory:
 
```{r}
getwd()
```
 
## Task 2

```{r}
epagas <- read.csv("EPAGAS-1.csv")
head(epagas)
```

## Task 3

 1. Make the object mpg, the number of miles per gallon vector.
 
```{r}
mpg = epagas$MPG
head(mpg)
```
 
 2. If $z_i=\frac{x_i-\bar{x}}{s_x}$, then $\bar{z}=0$ and $s^2_z=1$. Transform the mpg variable to z and verify these results.
 
```{r}
z = (mpg - mean(mpg)) / sd(mpg)
head(z)
print(paste0("mean = ", round(mean(z), digits = 4)), quote = FALSE)
print(paste0("variance = ", var(z)), quote = FALSE)
print(paste0("standard deviation = ", sd(z)), quote = FALSE)
```
 
 3. Using z, find the values of mpg that are possible outliers.
 
```{r}
mpg[abs(z) >= 2 & abs(z) <= 3]
```
 
 4. Using z, find the values of mpg that defined as outliers.
 
```{r}
mpg[abs(z) > 3]
```
 
 
 5. Using the lattice package construct a dotplot with colors, Red=outlier, Blue=possible outlier. (NB – read the instructions in the lab2.r file for installing the package)
 
```{r}
mycol = ifelse(abs(z) > 3, "red",
        ifelse(abs(z) >= 2 & abs(z) <= 3,"blue", "black")) 
library(lattice)
dotplot(mpg, col = mycol, main = "Plot of mpg values")
```

## Task 4

 1. Make a boxplot of the mpg variable. Make the box black. Put a notch where the median goes. Put a title on the graph. Make the plot horizontal.
```{r}
boxplot(mpg, main = "MPG Boxplot", xlab = "MPG", col = "black", horizontal = TRUE, notch = TRUE)
```
 
 2. Using Chebyshev’s theorem, $1 - \frac{1}{k^2}$, predict the proportion of data within 2 standard deviations of the mean of the mpg data.
```{r}
k = 2
cheb = 1 - 1 / (k ^ 2)
cheb
```
 
 3. Use R to calculate the exact proportion within 2 standard deviation of the mean.
```{r}
exacttwo = length(mpg[abs(z) < 2]) / length(mpg)
exacttwo
```
 
 
 4. Does Chebyshev agree with the data? `Yes`
 5. Now use the empirical rule, what proportion (according to the rule) of the data should be within 2 standard deviations of the mean? `95%`
 6. How well does it correspond? `Pretty well! 96% is very close to 95%.`
 7. Is the Empirical rule valid in this case? Why? `Yes. As shown by the histogram, it has the mound-shaped & symmetric distribution.`
  ```{r}
# X-axis grid
mpg2 <- seq(min(mpg), max(mpg), length = 40)
# Normal curve
curve <- dnorm(mpg2, mean = mean(mpg), sd = sd(mpg))
# Histogram
hist(mpg, prob = TRUE, col = "white",
     ylim = c(0, max(curve)),
     main = "Histogram of MPG with curve")
lines(mpg2, curve, col = 2, lwd = 2)
# Credit: https://r-charts.com/distribution/histogram-curves/
```
 
