---
title: "lab5"
author: "Joshua Wiseman"
date: "2023-02-12"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

## Working Directory
```{r}
getwd()
```


# Task 2

## mybin function
```{r}
mybin=function(iter=100,n=10, p=0.5){ 
  # make a matrix to hold the samples initially filled with NA's
  sam.mat=matrix(NA,nrow=n,ncol=iter, byrow=TRUE)
  
  #Make a vector to hold the number of successes in each trial
  succ=c()
  
  for(i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
    
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  
  #Make a barplot of the proportions
  iter.lab = paste0("iter = ", iter)
  n.lab = paste0("n = ",n)
  p.lab = paste0("p = ",p)
  lab = paste0(iter.lab, n.lab, p.lab, sep=", ")
  barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
  succ.tab/iter
}
```

### 100 iterations Plot
```{r}
bin100 = mybin(iter = 100, n = 10, p = 0.7)
```

### 200 iterations Plot
```{r}
bin200 = mybin(iter = 200, n = 10, p = 0.7)
```

### 500 iterations Plot
```{r}
bin500 = mybin(iter = 500, n = 10, p = 0.7)
```

### 1000 iterations Plot
```{r}
bin1000 = mybin(iter = 1000, n = 10, p = 0.7)
```

### 10000 iterations Plot
```{r}
bin10000 = mybin(iter = 10000, n = 10, p = 0.7)
```

## 10000 iteration table
```{r}
names(bin10000) = 0:10
bin10000
```

## Verify Values
```{r}
dbin.tab = round(dbinom(0:10, size = 10, prob = 0.7), 4)
names(dbin.tab) = 0:10
dbin.tab
```

# Task 3

## Sample without Replacement
```{r}
marbles = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0)
sample(marbles, size = 5, replace = FALSE, prob = NULL)
```

## Sample with Replacement
```{r}
sample(marbles, size = 5, replace = TRUE, prob = NULL)
```

## myhyper function
```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
  # make a matrix to hold the samples initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes over the trials
  succ=c()
  for( i in 1:iter){
  #Fill each column with a new sample
  sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
  #Calculate a statistic from the sample (this case it is the sum)
  succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
  succ.tab/iter
}
```

### 100 iterations Plot
```{r}
hyper100 = myhyper(iter=100, n=5, N=20, r=12)
```

### 200 iterations Plot
```{r}
hyper200 = myhyper(iter=200,n=5, N=20,r=12)
```

### 500 iterations Plot
```{r}
hyper500 = myhyper(iter=500,n=5, N=20,r=12)
```

### 1000 iterations Plot
```{r}
hyper1000 = myhyper(iter=1000,n=5, N=20,r=12)
```

### 10000 iterations Plot
```{r}
hyper10000 = myhyper(iter=10000,n=5, N=20,r=12)
```

## 10000 iteration table
```{r}
names(hyper10000) = 0:5
hyper10000
```

## Verify Values
```{r}
dhype = round(dhyper(x=0:5, m=12, n=8, k=5), 4)
names(dhype) = 0:5
dhype
```

# Task 4

## What does mysample() do?
```{r}
mysample=function(n, iter=10,time=0.5){
  for( i in 1:iter){
    #make a sample
    s=sample(1:10,n,replace=TRUE)
    # turn the sample into a factor
    sf=factor(s,levels=1:10)
    #make a barplot
    barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
    main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
    ylim=c(0,0.2)
    )
    
    #release the table
    Sys.sleep(time)
  }
}
```
The mysample function takes a vector n, number of iterations iter, and amount of time time. Basically it will sample 10 random samples from the vector, turn that sample to a factor, and then make barplot of that factor. Then the function will wait time time seconds and then loop iter-1 more times after the inital iteration.

## Run Line
```{r}
# Original Line
# mysample(n=1000, iter=30, time=1)
# Changed Line
mysample(n=1000, iter=1, time=0)
```

The line takes 30 seconds to run because it waits 1 second after each iteration, in this case 30 (30*1 = 30 seconds). But it also samples 10 random samples from the 1000 sided popultion vector.

# ???Task 5

## R Calculations

### Calculation 1
```{r}
choose(8,4)
```

### ???Calculation 2
```{r}

```

## More R Calculations

### More Calculations 1
```{r}
1 - pnbinom(q=10-3,size=3,prob=0.4)
```

### ???More Calculations 2
```{r}

```


# Task 6

## Load Custom Package
```{r}
bin = BestRPackage::abin(100, 10, 0.5)
bin
```



