---
title: "Lab1"
author: "Joshua Wiseman"
date: "2023-01-23"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

 Working directory:
 
```{r}
getwd()
```
 
## Task 2

```{r}
ddt <- read.csv("DDT-1.csv")
head(ddt)
```

## Task 3

### Base Questions
 1. What are the qualitative variables in “ddt”? `RIVER & SPECIES`
 2. What are the quantitative variables in “ddt”? `MILE, LENGTH, WEIGHT, & DDT`
 3. How many SPECIES are in the ddt data set? `3`
 4. Subset the ddt data set so that we see only the observations with SPECIES, LMBASS and WEIGHT bigger than 800gms
```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
```
 5. Subset the ddt data set so that we see only those observations where the RIVER is SCM and DDT bigger than 4.0
```{r}
with(ddt, ddt[DDT>4.0 & RIVER=="SCM",])
```

### Clicker Questions

 1. Find the mean length of fish in the ddt data frame!
```{r}
mean(ddt$LENGTH)
```
 
 2. What is the standard deviation of the weight of fish in the ddt data frame?
```{r}
sd(ddt$WEIGHT)
```
 
 3. Is this the plot of LENGTH vs WEIGHT from the ddt data frame? `No`
 4. If v=1:20 what is the last value of v/20? `1`
 
## Task 4

 Make a table of the rivers.
```{r}
rivers <- table(ddt$RIVER)
rivers
```
 Make a color coded barplot of the rivers.
```{r}
library(ggplot2)
barplot(rivers, col = rainbow(4))
```
 
 Make a table of the rivers crossed with the species of fish.
```{r}
rivsp <- table(ddt$RIVER, ddt$SPECIES)
rivsp
```
 Make a barplot of the rivers crossed with the species of fish.

```{r}
barplot(rivsp, beside=TRUE, col = rainbow(4))
with(ddt, legend(x = 1, y = 80, legend = ddt$RIVER[!(duplicated(ddt$RIVER))], bty = "n", fill = rainbow(4), cex = 0.8))
```

## Task 5

 Make a pie chart of Fish SPECIES
```{r}
spetab <- table(ddt$SPECIES)
pie(spetab,col=1:4)
```

 Make a pie chart of RIVERS
```{r}
rivtab <- table(ddt$RIVER)
pie(rivtab,col=1:4)
```

## Task 6
 
 Make a boxplot of the variable DDT
```{r}
layout(matrix(c(1,2,3),nr=1,nc=3))# 1 row 3 cols
with(ddt,boxplot(DDT,ylab="DDT",col="Red",notch=TRUE))
```

 Make a boxplot of the variable WEIGHT
```{r}
with(ddt,boxplot(WEIGHT,ylab="WEIGHT",col="Green",notch=TRUE))
```
 
 Make a boxplot of the variable LENGTH
```{r}
with(ddt,boxplot(LENGTH,ylab="LENGTH",col="Blue",notch=TRUE))
```
 
## Task 7

 Make a coplot of LENGTH Vs WEIGHT given RIVER
```{r}
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))
coplot(WEIGHT~LENGTH|RIVER,data=ddt, col = rivcol)
```

 Make a coplot of DDT Vs WEIGHT given SPECIES
```{r}
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(DDT~WEIGHT|SPECIES,data=ddt, col = fishcol)
```
 
## Task 8

 Graph 1
```{r}
library(ggplot2)
g1 <- ggplot(ddt)
g1 <- g1 + geom_boxplot(aes(x = SPECIES, y = WEIGHT, fill = RIVER)) + ggtitle("Joshua Wiseman")
g1
```

 Graph 2
```{r}
library(ggplot2)
g2 <- ggplot(ddt)
g2 <- g2 + geom_violin(aes(x = RIVER, y = LENGTH, fill = SPECIES)) + ggtitle("Joshua Wiseman")
g2
```

 Graph 3
```{r}
library(ggplot2)
g3 <- ggplot(ddt)
g3 <- g3 + geom_point(aes(x = WEIGHT, y = LENGTH, col = SPECIES)) + ggtitle("Joshua Wiseman")
g3
```
