---
title: "LAB10"
author: "Joshua Wiseman"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

## Working Directory
```{r}
getwd()
```

# Task 2 NOT DONE

## mymaxlik function
```{r}
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix – try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun) # A
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y >= max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y))) # B
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope should change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}
```

## A & B explination
A is the outer() function which is used to apply the lfun function to all possible combinations of x and param. z contains output lfun for each combination of x and param.

B finds the index of the maximum value of the y vector, which represents the log likelihood for each value of the parameter. which() returns the indices of the elements in y that are equal to the maximum value of y and max() returns the largest of these indices.

## 8 binom experiments
```{r}
y = c(3,3,4,3,4,5,5,4)
# mymaxlik(binom(8,,sum(y)),y)
```

# Task 3

## Likelihood of $\lambda$
```{r}
y = c(4,6,7,6,5)
mean(y)
```

## Graph Solution
```{r}
mymaxlik(y)
```

## Algebraic Expression
```{r}

```

## myNRML()
```{r}

```


# Task 4 Not done

# Task 5 Not done

# Task 6 Not done

# Task 7 Not done

# Task 8 Not done
